---
output:
  html_document: default
  pdf_document: default
---
First let's import the phenotype data 
```{r}
data<-read.csv(file="~/Dropbox/Projects/Carnea_Thanatosis/data/DeathFeignMaster_9APR.csv", header=TRUE)
names(data)[1]<-"animal"
names(data)[13]<-"fam"
head(data)
tail(data) 
```

Now let's import the pedegree data 
```{r}
ped<-as.data.frame(read.table(file="~/Dropbox/Projects/Carnea_Thanatosis/data/deathfeignped.txt", header=TRUE))
for(x in 1:3) ped[,x]<-as.factor(ped[,x])
head(ped)
tail (ped)
```

Now we load the required packages
```{r}
require(MCMCglmm)
```

Lets do a simple analysis using binary feign data with no additional fixed or random effects on all of the data. The Chi-squared prior which MCMCglmm tutorial recommends for binary data will be used.
```{r}
priora <- list(R = list(V = 1, fix = 1), 
              G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))
  
modela <- MCMCglmm(feign ~ 1, random = ~animal, 
                  family = "categorical", 
                  prior = priora, pedigree = ped, data = data, 
                  nitt = 100000, burnin = 10000, thin = 100, verbose=FALSE)
```

Now we look at how the analysis ran

```{r}
plot(modela$Sol)
autocorr.diag(modela$Sol)
effectiveSize(modela$Sol)

plot(modela$VCV)
autocorr.diag(modela$VCV)
effectiveSize(modela$VCV)

autocorr.diag(modela$Sol)
autocorr.diag(modela$VCV)

summary(modela)

```

Next calculate heritability 
```{r}
herita <- modela$VCV[, "animal"]/(modela$VCV[, "animal"] +
      modela$VCV[, "units"] + 1)
mean(herita)
HPDinterval(herita)

```

Now let's try to re-run analysis without any 1/2 sib data

```{r}
datanohalf <- data[!data$fam %in% c("", "E", "J", "L", "M", "N", "Q", "KK"),]
```

Here is a simple model with all 1/2 sib families excluded

```{r}
priorb <- list(R = list(V = 1, fix = 1), 
              G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))
  
modelb <- MCMCglmm(feign ~ 1, random = ~animal, 
                  family = "categorical", 
                  prior = priorb, pedigree = ped, data = datanohalf, 
                  nitt = 1000000, burnin = 2000, thin = 200, verbose=FALSE)
```

Now we look at how the analysis ran

```{r}
plot(modelb$Sol)
autocorr.diag(modelb$Sol)
effectiveSize(modelb$Sol)

plot(modelb$VCV)
autocorr.diag(modelb$VCV)
effectiveSize(modelb$VCV)

autocorr.diag(modelb$Sol)
autocorr.diag(modelb$VCV)

summary(modelb)

```


Next calculate heritability 
```{r}
heritb <- modelb$VCV[, "animal"]/(modelb$VCV[, "animal"] +
      modelb$VCV[, "units"] + 1)
mean(heritb)
HPDinterval(heritb)

```


Next we will try a model without any of the starved groups. 

```{r}
data$pdiet <- ifelse(is.na(data$crossf), data$diet, ifelse(data$crossf < 200, 1, ifelse(data$crossf < 300, 2, 3)))

datafedonly <- data[data$pdiet==1, ]
```

Now we do a simple analysis on only the fed group, but including 1/2 sibs

```{r}
priorc <- list(R = list(V = 1, fix = 1), 
              G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))
  
modelc <- MCMCglmm(feign ~ 1, random = ~animal, 
                  family = "categorical", 
                  prior = priorc, pedigree = ped, data = datafedonly, 
                  nitt = 110000, burnin = 10000, thin = 100, verbose=FALSE)
```


Now we look at how the analysis ran

```{r}
plot(modelc$Sol)
autocorr.diag(modelc$Sol)
effectiveSize(modelc$Sol)

plot(modelc$VCV)
autocorr.diag(modelc$VCV)
effectiveSize(modelc$VCV)

autocorr.diag(modelc$Sol)
autocorr.diag(modelc$VCV)

summary(modelc)

```



Next calculate heritability 
```{r}
heritc <- modelc$VCV[, "animal"]/(modelc$VCV[, "animal"] +
      modelc$VCV[, "units"] + 1)
mean(heritc)
HPDinterval(heritc)

```

Now let's make the model more complicated

First we will check for a maternal effect

```{r}
data$mom<-ped$mother
datafedonly <- data[data$pdiet==1, ]
tail(data)
```

Now we run the model with maternal effects
```{r}
priorm <- list(R = list(V = 1, fix = 1), 
                 G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1), 
                 G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

modelm<-MCMCglmm(feign ~ 1, random= ~animal+mom,
                   family="categorical", prior=priorm, pedigree=ped, data=datafedonly,
                   nitt = 110000, burnin = 10000, thin = 100, verbose=FALSE)
```

```{r}
plot(modelm$Sol)
autocorr.diag(modelm$Sol)
effectiveSize(modelm$Sol)

plot(modelm$VCV)
autocorr.diag(modelm$VCV)
effectiveSize(modelm$VCV)

autocorr.diag(modelm$Sol)
autocorr.diag(modelm$VCV)

anm <- summary(modelm)
DICm <- anm$DIC

```

Test effect of maternal effects when randomizing mothers on DIC
```{r}

nd <- datafedonly %>% 
  select(animal, feign, mom) 

reps <- 100

priorm <- list(R = list(V = 1, fix = 1), 
                 G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1), 
                 G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

out <- matrix(NA, nrow=reps, ncol=4)

for(i in 1:reps) {
  
  nd$mom <- sample(nd$mom, size=nrow(nd), replace=FALSE)
  
  modelm2<-MCMCglmm(feign ~ 1, random= ~animal+mom,
                   family="categorical", prior=priorm, pedigree=ped, data=nd,
                   nitt = 11000, burnin = 1000, thin = 10, verbose=FALSE)
  
  
 an_mean <- mean(modelm2$VCV[, "animal"])
 an_low <- quantile(modelm2$VCV[, "animal"], prob=0.025)
 an_high <- quantile(modelm2$VCV[, "animal"], prob=0.975)
 DIC <- modelm2$DIC
 
 out[i, ] <- cbind(an_mean, an_low, an_high, DIC)
  
}

save(sim_maternal, file="~/Dropbox/Projects/Carnea_Thanatosis/data/sim_maternal.RData")



```



```{r}
heritm <- modelm$VCV[, "animal"]/(modelm$VCV[, "animal"] +
      modelm$VCV[, "units"] + 1)
mean(heritm)
HPDinterval(heritm)

```

We are trying to make a ped with just fed families
```{r}
fedped<-as.data.frame(read.table(file="~/Dropbox/Projects/Carnea_Thanatosis/data/deathfeignpedfed.txt", header=TRUE))
for(x in 1:3) ped[,x]<-as.factor(ped[,x])
head(fedped)
tail(fedped)
```


We will now construct a dominance matrix and add a column to data called dom that is identical to animal... the package needs this


```{r}
require(nadiv)
datafedonly$dom<-datafedonly$animal
listD<-makeD(fedped)
#Dinv<-as.matrix(listD$Dinv)
Dinv<-listD$Dinv


rownames(Dinv)<- datafedonly$dom
```



```{r}
priord <- list(R = list(V = 1, fix = 1), 
          G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1), 
          G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

modeld<-MCMCglmm(feign ~ 1, random= ~animal+dom, ginverse=list(dom=Dinv),
                family="categorical", prior=priord, pedigree=ped, 
                data=datafedonly,nitt=100000,burnin=2000,thin=100, verbose=FALSE)
```


```{r}
plot(modeld$Sol)
autocorr.diag(modeld$Sol)
effectiveSize(modeld$Sol)

plot(modeld$VCV)
autocorr.diag(modeld$VCV)    
effectiveSize(modeld$VCV)

autocorr.diag(modeld$Sol)
autocorr.diag(modeld$VCV)

summary(modeld)

```

Test effect of dominance effects when randomizing pedigree on DIC
```{r}


nd <- datafedonly %>% 
  select(animal, feign, dom) 

pd <- fedped
listD<-makeD(pd)
Dinv<-listD$Dinv
rownames(Dinv)<- nd$dom

reps <- 10

priord <- list(R = list(V = 1, fix = 1), 
          G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1), 
          G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

sim_dom <- matrix(NA, nrow=reps, ncol=4)

for(i in 1:reps) {
  
  pd$mother <- sample(pd$mother, size=nrow(pd))
  pd$father <- sample(pd$father, size=nrow(pd))
  
  listD<-makeD(pd)
  Dinv<-listD$Dinv
  
  modeld2<-MCMCglmm(feign ~ 1, random= ~animal+dom, ginverse=list(dom=Dinv),
                family="categorical", prior=priord, pedigree=ped, 
                data=nd,nitt=11000,burnin=1000,thin=10, verbose=FALSE)
  
  
 an_mean <- mean(modeld2$VCV[, "animal"])
 an_low <- quantile(modeld2$VCV[, "animal"], prob=0.025)
 an_high <- quantile(modeld2$VCV[, "animal"], prob=0.975)
 DIC <- modeld2$DIC
 
 sim_dom[i, ] <- cbind(an_mean, an_low, an_high, DIC)
  
}

save(sim_maternal, file="~/Dropbox/Projects/Carnea_Thanatosis/data/sim_maternal.RData")



```





```{r}
heritd <- modeld$VCV[, "animal"]/(modeld$VCV[, "animal"] +
      modeld$VCV[, "units"] + 1)
mean(heritd)
HPDinterval(heritd)

```


